Narrative

### 2023-05-21 manuscript/introduction.qmd
intro: added references to selective reporting part

One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated (e.g., REF1, REF2) due to factors such as publication bias (REF1) and _p_-hacking (REF2), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015). In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.

@@

It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the positive bias (e.g., Borenstein, 2009, p. 342; Card, 2015, p. 126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020). However, given the value of accurate  heterogeneity estimates for many research outcomes in psychology and that the negative bias could often be 20% or more, we consider it important for meta-analysts to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability. This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies. Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates when meta-analyzing primary studies with imperfect measurement reliabilities.

>>>>

One reason that few meta-analysts correct for unreliability could be that correcting for unreliability is generally controversial and has been so since it was proposed (for arguments spanning the last century, see Table 2 of LeBreton et al., 2017). On the one hand, correcting for unreliability is seen by some as conceptually problematic because it inflates effect size estimates to match a hypothetical and unachievable scenario of perfect measurement (e.g., Seymour, 1988; LeBreton et al., 2017). Systematically correcting for unreliability may also be expected to lead to overestimation of effect sizes, because reliability estimates are lower bound estimates (i.e., a reliability estimate of 0.8 implies the reliability is between 0.8 - 1). Even worse, the most commonly reported estimate of reliability in psychology (Chronbach's alpha; Flake et al., 2017) as well as improvements thereof tend to underestimate the lower bound of reliability (REF Klaas). Moreover, from a pragmatic meta-analytic perspective, effect sizes appear to often be overestimated in psychology (e.g., Kvarven et al., 2019; Schäfer & Scharz, 2019; Scheel et al., 2020) due to selective reporting based on the significance of outcomes (Simmons et al., 2011) and publication bias (REF?), so correcting for unreliability may inflate meta-analytic estimates even further and lead to estimates further from their true values rather than closer. On the other hand, researchers in psychology are typically interested in latent constructs rather than observed measures. As such, neglecting measurement artifacts means that computed estimates do not correspond to the entity of interest. From this perspective, correcting for unreliability is desirable even if doing so is challenging (e.g., Oswald et al., 2015; Hunter & Schmidt, 2015). In the end, whether correcting for unreliability or not, authors and consumers of meta-analyses need to be aware of how this decision affects heterogeneity estimates.

@@

It is difficult to predict the total effect of the negative and positive bias in heterogeneity due to unreliability, and existing literature tends to focus on the positive bias (e.g., Borenstein, 2009, p. 342; Card, 2015, p. 126; Hunter & Schmidt, 2015; Wiernick & Dahlke, 2020). However, given the value of accurate  heterogeneity estimates for many research outcomes in psychology and that the negative bias could often be 20% or more, we consider it important for meta-analysts to have insight into the total bias in heterogeneity estimates that should be expected due to unreliability. This depends on average effect size, true heterogeneity, mean reliability across studies, the variability in reliabilities, and sampling variance within studies. Hence, we perform a Monte-Carlo simulation study to explore the expected bias in heterogeneity estimates due to unreliability in primary studies.
